---
title: "Regimens"
always_allow_html: yes
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_vignette:
    toc: yes
    toc_depth: 3
    vignette: >
      %\VignetteIndexEntry{Regimens}
      %\VignetteEncoding{UTF-8}
      %\VignetteEngine{knitr::rmarkdown}
  pdf_document:
    toc: yes
---

>A regimen is a plan, or course of action such as a diet, exercise or medical treatment. A low-salt diet is a regimen. A course of penicillin is a regimen, and there are many chemotherapy regimens in the treatment of cancer.

[*Wikipedia, 2025*](https://en.wikipedia.org/wiki/Regimen)

Regimans are a set of a actions that combined form a form of therapy. TreatmentPatterns can handle regimens in two ways:

1. Regimen as a single event
2. Identifying regimens from pathways

## Regimen as a single event
This is arguably the most straightforward way to deal with regimens. You can pre-specify a cohort using ATLAS or `Capr` to the regimens as a single cohort entry. TreatmentPatterns would then pick up these entries while building the pathways. Let's say we have two singular events A and C, and a regimen; B:

```
Event A -> Regimen B -> Event C
```

## Identifying regimens from pathways
Another way we can interact with regimens through TreatmentPatterns is by looking for relevant events during some period.

As an example we can take the ["7+3"](https://en.wikipedia.org/wiki/7%2B3_(chemotherapy)) regimen, which refers to a regimen of two treatments: Cytarabine; and Daunocrubicin, (or Idarubicin, or Mitoxantrone). On day 1 through 7 Cytarabine is given, and on day 1 to 3 Daunocrubicin (or alternative). Theoretically, this gives us an overlap between the two individual treatments of 3 days.

| Treatment | Day 1 | Day 2 | Day 3 | Day 4 | Day 5 | Day 6 | Day 7 |
| --- | :---: | :---: | :---: | :---: | :---: | :---: | :---: |
| Cytarabine |   X   |   X   |   X   |   X   |   X   |   X   |   X   |
| Daunocrubicin |    X   |   X   |   X   | | | | |

According to the article above, both treatments are IV's, they most likeley show up in the `drug_exposure` table as re-occurring records, starting and ending on the same day. **This can (and will) vary per database**. In the `drug_era` table they can however be concatinated with other drug exposures of the same ingredient. Because for this specific regimen the drug exposures are IV's, **carefull phenotyping and concept set curation are key** to eliminate most of the noise from other (irrelevant) drug exposures. This makes databases that do not have more detailed descendants of ingredient drug exposures not well suited for identifying this regimen.

The sensitivity to noise in the concept sets can vary between regimens. For a regimen where a drug can be administered in multiple ways (injection, oral, intravenous, etc.) require more pruning than a drug that is only administered in one way. If this pruning is not done correctly, you will most likely heavily over estimate the amount of people that receive your regimen of interest.

TreatmentPatterns uses `event` and  `target` cohorts. If you're not familiar with this terminology, please read the Computing Pathways vignette first.

We'll do something quite unusual here. We will use **one** of our regimen treatments as **both** a `target` and `event` cohort. If you expect the regimen to occur during a condition from the `condition_occurrence` table, this is the cohort where you would embed that condition in the *inclusion criteria*. The *Cohort Entry Event* is the treatment in the regimen that you expect to occur **first**. If two or more treatments can occur first, you can just pick one, for which you want to act as the target cohort.

### In code
With our cohorts (with carefully curated concept sets) in hand, we can generate the cohorts in a cohort table, and specify our parameters in `TreatmentPatterns`.

For this specific regimen, we know that the overlap between both treatments *must* be 3 days. So we set our `combinationWindow = 3`. One thing to consider here is that this setting requires *at least*  3 days over overlap, and not exactly 3 days. We can set `minEraDuration = 1` and `minPostCombinationDuration = 1` to keep all events that last at least one day, as our events of interest should last longer than that.

These settings will vary from regimen to regimen, ideally they should resemble the what happens in the real world. You can add a sensitivity analysis with variants to see what seems to work best.

We will setup some dummy data first:
```{r}
library(DBI)
library(duckdb)
library(CDMConnector)
library(TreatmentPatterns)

# Connect to our Eunomia CDM
con <- DBI::dbConnect(
  drv = duckdb::duckdb(),
  dbdir = CDMConnector::eunomiaDir()
)

# Dummy cohort table
cohort_table <- dplyr::tribble(
  ~cohort_definition_id, ~subject_id, ~cohort_start_date,    ~cohort_end_date,
  1,                     5,           as.Date("2014-01-01"), as.Date("2014-01-07"),
  2,                     5,           as.Date("2014-01-01"), as.Date("2014-01-03")
)

# Copy cohort table to the database
dplyr::copy_to(con, cohort_table, overwrite = TRUE)

# Create a cdm reference object
cdm <- CDMConnector::cdmFromCon(
  con = con,
  cdmSchema = "main",
  writeSchema = "main",
  cohortTables = "cohort_table"
)
```

```{r}
cohorts <- data.frame(
    # We list Cytarabine twice as cohort 1, as mentioned earlier
  cohortId = c(1, 1, 2),
  cohortName = c("Cytarabine", "Cytarabine", "Daunocrubicin"),
  type = c("target", "event", "event")
)

andromeda <- TreatmentPatterns::computePathways(
  cohorts = cohorts,
  cohortTableName = "cohort_table",
  cdm = cdm,

  # Our parameters we specified earlier
  minEraDuration = 1,
  combinationWindow = 3,
  minPostCombinationDuration = 1
)

result <- TreatmentPatterns::export(andromeda, minCellCount = 1)

result$treatment_pathways
```

Sometimes treatments in your regimen will start after each other. What can happen is that you can lose events when you're trying to build the pathways, because your target cohort is also an event cohort. As an example lets say the Daunocrubicin treatment starts on day 8 and ends on day 7:

| Treatment | Day 1 | Day 2 | Day 3 | Day 4 | Day 5 | Day 6 | Day 7 | Day 8 | Day 9 | Day 10 |
| --- | :---: | :---: | :---: | :---: | :---: | :---: | :---: | :---: | :---: | :---: |
| Cytarabine |   X   |   X   |   X   |   X   |   X   |   X   |   X   |
| Daunocrubicin | | | | | | | | X | X | X |

```{r}
library(DBI)
library(duckdb)
library(CDMConnector)
library(TreatmentPatterns)

# Connect to our Eunomia CDM
con <- DBI::dbConnect(
  drv = duckdb::duckdb(),
  dbdir = CDMConnector::eunomiaDir()
)

# Dummy cohort table
cohort_table <- dplyr::tribble(
  ~cohort_definition_id, ~subject_id, ~cohort_start_date,    ~cohort_end_date,
  1,                     5,           as.Date("2014-01-01"), as.Date("2014-01-07"),
  2,                     5,           as.Date("2014-01-08"), as.Date("2014-01-10")
)

# Copy cohort table to the database
dplyr::copy_to(con, cohort_table, overwrite = TRUE)

# Create a cdm reference object
cdm <- CDMConnector::cdmFromCon(
  con = con,
  cdmSchema = "main",
  writeSchema = "main",
  cohortTables = "cohort_table"
)
```

```{r}
cohorts <- data.frame(
    # We list Cytarabine twice as cohort 1, as mentioned earlier
  cohortId = c(1, 1, 2),
  cohortName = c("Cytarabine", "Cytarabine", "Daunocrubicin"),
  type = c("target", "event", "event")
)

andromeda <- TreatmentPatterns::computePathways(
  cohorts = cohorts,
  cohortTableName = "cohort_table",
  cdm = cdm,

  # Our parameters we specified earlier
  minEraDuration = 1,
  combinationWindow = 3,
  minPostCombinationDuration = 1
)

result <- TreatmentPatterns::export(andromeda, minCellCount = 1)

result$treatment_pathways
```

What we see is that we are now only left with the Cytarabine treatment. But we expected a pathway of `Cytarabine-Daunocrubicin`. We can then use the windowing parameters to alter the target window of our analysis.

We can set the `startAnchor = "startDate"` to anchor on the start date of our target cohort. We can then set `windowStart = 0` to offset the start by 0 days, meaning it will just be the index date. We can then offset our end date by setting `windowEnd = 10` to extend the target window to 7 days, rather than the 3 days we actually have in the data:

```{r}
cohorts <- data.frame(
  # Now Daunocrubicin is the target cohort
  cohortId = c(1, 1, 2),
  cohortName = c("Cytarabine", "Cytarabine", "Daunocrubicin"),
  type = c("target", "event", "event")
)

andromeda <- TreatmentPatterns::computePathways(
  cohorts = cohorts,
  cohortTableName = "cohort_table",
  cdm = cdm,

  # Our parameters we specified earlier
  minEraDuration = 1,
  combinationWindow = 3,
  minPostCombinationDuration = 1,

  startAnchor = "startDate",
  windowStart = 0,
  windowEnd = 10
)

result <- TreatmentPatterns::export(andromeda, minCellCount = 1)

result$treatment_pathways
```

